Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-11-30 main.tex
State that 24 participants of 26 were availabe to analyze.

Due to data quality issues that prevented time synchronization of the
measurements, we excluded one participant from the 6~\si{\kph} dataset and one
from the 10~\si{\kph} dataset, leaving 10 and 14 included participants with 40
perturbations each. Before fitting the model, we scale each independent
variable \(x_{ij}^k\) such that they have a mean of zero and a standard
deviation of one by cluster-mean centering with clusters being an individual
participant, as recommended by \cite{Enders2007}. The clusters are chosen as
all data from an individual participant because we are only interested in the
association between the state of the balance assist system and the outcome of
the perturbation for that participant, not how participants perform relative to
each other. We expected there to be variation between participants in how well
they are able to resist the perturbation; however, this was not true. If a
random effect is included in the model for each participant, the intraclass
correlation (ICC) is less than 3\% for both the 6~\si{\kph} and 10~\si{\kph}
datasets, so we utilize a simple single-level logistic regression model without
the random effect instead of a multilevel model. This tracks with the
independent data collected in \cite{Haitjema2023a} which also had negligible
participant variation. This left us with angular impulse, perturbation order,
balance assist state, and roll \& steer angles at the time of perturbation as
independent variables. We also include interaction effects between the balance
assist state and the other four variables. We divide the analysis into two
separate model evaluations, one for the 6~\si{\kph}, \(\kappa=5.2\) trials and
one for the 10~\si{\kph}, \(\kappa=3.9\) trials and we evaluate our hypothesis
for each set of data. Statistics were computed with R version
4.1.3~\cite{RCoreTeam2021} and the lme4 version 1.1~\cite{Bates2015} package.

>>>>

Due to data quality issues that prevented time synchronization of the
measurements, we excluded one participant from the 6~\si{\kph} dataset and one
from the 10~\si{\kph} dataset, leaving 24 participants for the analysis. Before
fitting the model, we scale each independent variable \(x_{ij}^k\) such that
they have a mean of zero and a standard deviation of one by cluster-mean
centering with clusters being an individual participant, as recommended by
\cite{Enders2007}. The clusters are chosen as all data from an individual
participant because we are only interested in the association between the state
of the balance assist system and the outcome of the perturbation for that
participant, not how participants perform relative to each other. We expected
there to be variation between participants in how well they are able to resist
the perturbation; however, this was not true. If a random effect is included in
the model for each participant, the intraclass correlation (ICC) is less than
3\% for both the 6~\si{\kph} and 10~\si{\kph} datasets, so we utilize a simple
single-level logistic regression model without the random effect instead of a
multilevel model. This tracks with the independent data collected in
\cite{Haitjema2023a} which also had negligible participant variation. This left
us with angular impulse, perturbation order, balance assist state, and roll \&
steer angles at the time of perturbation as independent variables. We also
include interaction effects between the balance assist state and the other four
variables. We divide the analysis into two separate model evaluations, one for
the 6~\si{\kph}, \(\kappa=5.2\) trials and one for the 10~\si{\kph},
\(\kappa=3.9\) trials and we evaluate our hypothesis for each set of data.
Statistics were computed with R version 4.1.3~\cite{RCoreTeam2021} and the lme4
version 1.1~\cite{Bates2015} package.
